\documentclass{article}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{centernot}
\author{Mostafa Hassanein}
\title{
  MTH-684 Logic \\
  Assignment (2): Propositional Logic}
\date{2 October 2024}
\begin{document}

\maketitle
\newpage

\section*{2-3}

\textbf{\underline{Note: }} I've only managed to edit part (a).


\subsection*{a.}
\begin{proof}
  $ $

  Let $A$ be a truth assignment function and $[[.]]^A$ be its corresponding interpretation function under which $(\Gamma_1 \models \phi) = (\Gamma_2 \models \psi) = \top$.
  \newline

  We now want to show that $(\Gamma_1 \cup \Gamma_2 \models \phi \land \psi) = \top$ under the same interpretation function $[[.]]^A$. 
    
  So suppose that all formulas in $\Gamma_1 \cup \Gamma_2$ interpret to $\top$ under $[[.]]^A$. 

  Because all formulas in $\Gamma_1$ are $\top$ and $\Gamma_1 \models \phi$, we have $\phi = \top$.

  And because all formulas in $\Gamma_2$ are $\top$ and $\Gamma_2 \models \psi$, we have $\psi = \top$.

  Finally, from the definition of logical conjunction: $\phi \land \psi = \top \iff \phi = \psi = \top$, we conclude that $\phi \land \psi = \top$.

\end{proof}

\subsection*{b.}
\begin{proof}
  $ $

  Let $T_4 = T_1 \cup T_2 \cup T_3$.

  Suppose all WFFs in $T_4$ are true, then all WFFs in $T_1$, $T_2$, and $T_3$ are true.
  
  From $T_1 \models \phi$, we conclude that $\phi \lor \psi$ is true. We have 3 cases:
  \newline

  \underline{1. $\phi = \top \text{ and } \psi = \bot$:}
  
  In this case, all WFFs in $T_2 \cup \{ \phi \}$ are true.
  
  From $T_2 \cup \{ \phi \} \models \zeta$, we conclude that $\zeta$ is true.
  \newline

  \underline{2. $\phi = \bot \text{ and } \psi = \top$:}
  
  In this case, all WFFs in $T_3 \cup \{ \psi \}$ are true.
  
  From $T_3 \cup \{ \psi \} \models \zeta$, we conclude that $\zeta$ is true.
  \newline

  \underline{3. $\phi = \top \text{ and } \psi = \top$:}
  
  Using either case 1 or 2, it is clear that $\zeta$ is true.
  \newline

  Therefore, we conclude that $T_4 = T_1 \cup T_2 \cup T_3 \models \zeta$.  

\end{proof}

\subsection*{c.}
\begin{proof}
  $ $

  We want to show that $T_1 \models (\phi \implies \psi)$ is true, so we assume $T_1$ and attempt to show that $\phi \implies \psi$ holds.

  Next, to show that $\phi \implies \psi$, we assume $\phi$ and attempt to show that $\psi$ holds.

  Putting our assumptions together: $T_1$ and $\phi$ are true. Therefore, all WFFs in $T_1 \cup \{ \phi \}$ are true.

  From $T_1 \cup \{ \phi \} \models \psi$, we conclude that $\psi$ is true.

  Therefore, $T_1 \models (\phi \implies \psi)$ as desired.

\end{proof}

\subsection*{d.}
\begin{proof}(By Contradiction)
  $ $

  Suppose $T_1 \cup T_2$ is true.

  For the sake of contradiction, suppose that $\phi$ is true.

  From $T_1 \cup \{ \phi \} \models \psi$, we conclude that $\psi$ is true.

  From $T_2 \cup \{ \phi \} \models \lnot \psi$, we conclude that $\lnot \psi$ is true.

  But then we have: $\psi \land \lnot \psi$ is true. This is a contradiction.

  Therefore, we must conclude that: $T_1 \cup T_2 \models \lnot \phi$.

\end{proof}

\end{document}