\documentclass{article}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{centernot}
\usepackage{makecell}
\author{Mostafa Hassanein}
\title{
  GEN-600 Technical Writing \\
  Assignment (2)}
\date{11 March 2025}
\begin{document}

\maketitle
\newpage

\section*{Paragraph Analysis}

\begin{center}
  \underline{Paragraph 1}
\end{center}

Writing is not simply speech written down on paper.  Learning to write is not just a natural extension of learning to speak a language.  We learned to speak our first language at home without systematic instruction; whereas, most of us had to be taught in school how to write that same language.  Many adult native speakers of a language find writing difficult. A speaker speaks to a listener who is right there, nodding or frowning, or interrupting or questioning.  For the writer, the reader's response is either delayed or nonexistent.
\newline


\textbf{Topic Statement:} Writing is different from speaking. 
\newline
\indent
\textbf{Topic:} Writing.
\newline
\indent
\textbf{Comment(s):} is not simply speech written down on paper.
\newline
\indent
\textbf{Grammatical Analysis:}
\newline
\newline

\begin{tabular}{c | c c c c c c c}
  Sentence & [1] & [2] & [3] & [4] & [5] & [6] \\
  \cline{1-7}
  Writing is not simplfy speech written down on paper. & [1] & [2] & [3] & [4] & [5] & [6] \\
  \hline
  Learning to write is
not just a natural extension of learning to speak a language. & [2] & [4] & [6] & [1] & [3] & [5] \\
\hline
\makecell{We learned to
speak our first language at home without systematic instruction; \\ whereas, most
of us had to be taught in school how to write that same language.} & [3] & [6] & [2] & [5] & [1] & [4] \\
  \hline
  {[4]} & [4] & [1] & [5] & [2] & [6] & [3] \\
  \hline
  {[5]} & [5] & [3] & [1] & [6] & [4] & [2] \\
  \hline
  {[6]} & [6] & [5] & [4] & [3] & [2] & [1] \\
  \hline
\end{tabular}


\newpage




\newpage
\section*{2. Action Verbs}
1. Optimize, minimize, maximimize.
2. Speed up.
3. Accelerate.
4. Enhance.
5. Stabilize.
6. Streamline.
7. Synchronize.
8. Standardize.
9. destroy.
10. distill.
11. Cement.
12. Generate.
13. illuminate.
14. illustrate.
15. identify.
16. Kick.
17. Map
18. Mine.
19. Nurture.
20. Prove.
21. Trim.
22. Trap.
23. Test.
24. 

\newpage
\section*{3. Write a paragraph about 'Your Research Work'}


\end{document}